---
title: "Analysis of Clk856 TrpA1 fruFLP activiy data"
output: html_notebook
---

Analysis of data from ShinyR-DAM using R scripts from: Cichewicz, K., Hirsh, J. ShinyR-DAM: a program analyzing Drosophila activity, sleep and circadian rhythms. Commun Biol 1, 25 (2018). https://doi.org/10.1038/s42003-018-0031-9 
```{r}
library(knitr)
library(ggplot2)
library(gridExtra)
library(grid)
library(dplyr)
library(plyr)
library(zoo)
library(gtools)
library(scales)
library(gridExtra)
library(data.table)
```


```{r}
Average_individual_daily_locomotor_activity<-read.csv("/Volumes/NO NAME/LDclkact/Average_individual_daily_locomotor_activity.csv")

knitr::kable(head(Average_individual_daily_locomotor_activity))
```


Order list 
```{r}

Average_individual_daily_locomotor_activity$Condition <-factor(Average_individual_daily_locomotor_activity$Condition,levels = c("CS_M","CS_F", "Trp_Fru_FLP_M","Trp_Fru_FLP_F","CLK_TRPA1_M","CLK_TRPA1_F"))
```

#Supplemental Figure 14A
```{r}

Average_individual_daily_locomotor_activity <-filter(Average_individual_daily_locomotor_activity, Light_cycle == 'LD')
#Box plot of individual locomotor activity in LD
box_plot_LD<- ggplot(na.omit(Average_individual_daily_locomotor_activity), aes(x=Condition, y=mean_value, fill=Condition)) +
geom_boxplot(alpha=0.7)+
geom_point()+
labs(title= "Activity per day in LD")+
labs(y="Locomotor activity [counts/day]", x="") + #adds/removes axis lables
theme_bw()+
theme(legend.position="none")+
theme(legend.text=element_text(size=16))+
theme(legend.position="none")+
theme(axis.text.x=element_text(angle=50, vjust=0.9, hjust=1, size=15, face="bold"))+
theme(plot.title = element_text(size = rel(2), hjust=0.5))+
theme(axis.text.y=element_text(vjust=0.9, hjust=1, size=11, face="bold"))+
theme(axis.title.y = element_text(size=14))

box_plot_LD
```

```{r}
mat_keep_rows <- c("CS_M", "Trp_Fru_FLP_M", "CLK_TRPA1_M") 
MaleSubset <- subset(Average_individual_daily_locomotor_activity_LD, Condition %in% mat_keep_rows)
mat_keep_rows <- c("CS_F", "Trp_Fru_FLP_F", "CLK_TRPA1_F")
FemaleSubset <- subset(Average_individual_daily_locomotor_activity_LD, Condition %in% mat_keep_rows)

```


```{r}
model <- res.aov <- aov(mean_value ~ Condition,data=Average_individual_daily_locomotor_activity)
summary(model)
TukeyHSD(model)
```


#Supplemental Figure 14G
```{r}
Daytime_nighttime_activity<-read.csv("/Volumes/NO NAME/LDclkact/Daytime_nighttime_activity_in_LD-1.csv") 
knitr::kable(head(Daytime_nighttime_activity))
```
#order the plot
```{r}
Daytime_nighttime_activity$Condition <-factor(Daytime_nighttime_activity$Condition,levels = c("CS_M","CS_F", "Trp_Fru_FLP_M","Trp_Fru_FLP_F","CLK_TRPA1_M","CLK_TRPA1_F"))

```
 
```{r}
Daytime_nighttime_activity_plot <- ggplot(Daytime_nighttime_activity, aes(factor(Condition), Mean, fill = Light_status)) +
geom_bar(stat="identity", position = "dodge", colour="black") +
geom_errorbar(data=Daytime_nighttime_activity, aes(ymin=Mean - SEM, ymax=Mean + SEM),
position = position_dodge(0.9), width=0.1) +
scale_fill_manual(values=c("#E8E823", "#0472CC"))+
theme_bw()+
theme(legend.text=element_text(size=16))+
theme(axis.text.x=element_text(angle=50, vjust=0.9, hjust=1, size=13, face="bold"))+
labs(title= "Daytime vs nighttime activity in LD", size= 14)+
theme(plot.title = element_text(size = rel(2), hjust=0.5))+
theme(axis.text.y=element_text(vjust=0.9, hjust=1, size=11, face="bold"))+
theme(axis.title.y = element_text(size=14))+
theme(legend.title=element_blank())+
theme(axis.title.x = element_blank())+
labs(x="", y="Locomotor activity [counts/day]")

Daytime_nighttime_activity_plot
```
 
 
Actograms
#Supplemental Figure 14H-I
```{r}
Mean_and_median_actogram_data <- read.csv("/Volumes/NO NAME/LDclkact/Mean_and_median_actogram_data.csv") #Reads a csv file
knitr::kable(head(Mean_and_median_actogram_data)) #Displays the first 6 rows of a data frame
```

# Setting plot settings as variables allows to easily edit the plot without searching for parameters deep in the function
```{r}
actogram_bin <- 5 # Plot bin size [min]
data_recording_frequency <- 1 # Data acquisition frequency [min]
Double_Single <- 'SP' # Select 'SP' for single plotted actograms, and 'DP' for double plotted
ac_max_counts <- 5 # Max value of counts displayed on an actogram
data_freq <- 1440 # Number of data records in a day, Set to 1440/5 if your DAM system saves counts every 5 min.
mean_or_median_column <- "binned_mean" # Set to "binned_median" for median actogram
```

Make an object the same name as they use in the example for ease.
```{r}
mmf<-Mean_and_median_actogram_data
mmf <- filter(Mean_and_median_actogram_data, Condition %in%unique(Mean_and_median_actogram_data$Condition)[1:6])
```

```{r fig.height=60, fig.width=20}
# Function defining mean actogram plot
mean_actogram <- ({function(x){
# Binning actogram mean values - deccreases data resolution for plotting
y <- filter(mmf, Condition==x)
binning_value <- actogram_bin / data_recording_frequency #Generated a value for binning
y$binned_mean <- rep(rollapply(y$mean, width = binning_value, by = binning_value,
FUN = mean, align = "left"), each=binning_value)
y$binned_median <- rep(rollapply(y$median, width = binning_value, by = binning_value,
FUN = mean, align = "left"), each=binning_value)
# Uses a plotting function for a single plotted actogram if Double_Single <- 'SP' is set by user
if (Double_Single == 'SP') {
ggplot(y, aes(Dec_time, y=get(mean_or_median_column), ymax=get(mean_or_median_column),
ymin=min(get(mean_or_median_column)))) +
geom_ribbon() +
facet_grid(date ~ .)+
geom_ribbon(fill="#0D226E") +
labs(title= x, x= "", y = "Counts/recording frequency")+
theme_bw()+
scale_x_continuous(breaks = c(0, 360,720, 1080, (1440-data_recording_frequency)),
labels=c("0 h", "6 h", "12 h", "18 h", "0 h")) +
theme(axis.text=element_text(size=14))+
theme(text = element_text(size=16))+
theme(plot.title = element_text(size = rel(2), hjust=0.5))+
coord_cartesian(ylim=c(0,as.numeric(ac_max_counts)))
} else {
# Uses a plotting function for a double plotted actogram if Double_Single <- 'DP' is set by the user
# Well, actually any value other than 'SP' would do it
# Thsese 3 lines double the actogram data
b <- arrange(y, date, Dec_time)
qqq <- lapply(unique(b$date), function(w) filter(b, date==w))
zr1 <- do.call("rbind", replicate(2, qqq, simplify = T))
# Generates a doubled dec time X scale
zr1$Dec_time_double <- c(rep(1:data_freq, length(unique(zr1$date))),
rep((data_freq+1):(data_freq*2), length(unique(zr1$date))))*data_recording_frequency
zr2 <- arrange(zr1, date, Dec_time_double)
zr3 <- zr2[(data_freq+1):nrow(zr2),] #drops the first repeated day
zr3 <- arrange(zr3, date, Dec_time_double)

#Generates another doubled dec time X scale after dropping the 1st day
zr3$Dec_time_double2 <- ((c(rep(1:(data_freq*2), length(unique(zr3$date))))[1:nrow(zr3)])* data_recording_frequency)

ggplot(zr3, aes(Dec_time_double2, y=get(mean_or_median_column), ymax=get(mean_or_median_column),
ymin=min(get(mean_or_median_column)))) +
theme_bw()+
facet_grid(date2 ~ .)+
geom_ribbon(fill="#0D226E") +
labs(title= x, x= "", y = "Number of counts/recording frequency")+
scale_x_continuous(breaks = c(1, 360,720, 1080, 1440, 1800, 2160, 2520, 2880),
labels=c("0h", "6 h", "12 h", "18 h", "0 h", "6 h", "12 h", "18 h", "0 h"))+
coord_cartesian(xlim=c(1,2880))+
theme(plot.title = element_text(size = rel(2), hjust=0.5))+
theme(axis.text=element_text(size=14))+
theme(text = element_text(size=16))+
coord_cartesian(ylim=c(0,as.numeric(ac_max_counts)))
}
}
})
# Assembles mean actograms using marrangeGrob
plot<-marrangeGrob(lapply(unique(mmf$Condition), function(x) FUN=mean_actogram(x)),
ncol=1, nrow = length(unique(mmf$Condition)), top ="")

ggsave("multipage.pdf", plot, width = 10, height = 80, units = "in",limitsize = FALSE)


```


DD data actograms
#Supplemental Figure 15A-B
```{r}
Mean_and_median_actogram_data <- read.csv("/Volumes/NO NAME/Clk856DDN/Mean_and_median_actogram_data(1).csv") #Reads a csv file
knitr::kable(head(Mean_and_median_actogram_data)) #Displays the first 6 rows of a data frame
```

```{r}
actogram_bin <- 5 # Plot bin size [min]
data_recording_frequency <- 1 # Data acquisition frequency [min]
Double_Single <- 'SP' # Select 'SP' for single plotted actograms, and 'DP' for double plotted
ac_max_counts <- 5 # Max value of counts displayed on an actogram
data_freq <- 1440 # Number of data records in a day, Set to 1440/5 if your DAM system saves counts every 5 min.
mean_or_median_column <- "binned_mean" # Set to "binned_median" for median actogram
```

Make an object the same name as they use in the example for ease.
```{r}
mmf<-Mean_and_median_actogram_data
mmf <- filter(Mean_and_median_actogram_data, Condition %in%unique(Mean_and_median_actogram_data$Condition)[1:6])
```

```{r fig.height=60, fig.width=20}
# Function defining mean actogram plot
mean_actogram <- ({function(x){
# Binning actogram mean values - deccreases data resolution for plotting
y <- filter(mmf, Condition==x)
binning_value <- actogram_bin / data_recording_frequency #Generated a value for binning
y$binned_mean <- rep(rollapply(y$mean, width = binning_value, by = binning_value,
FUN = mean, align = "left"), each=binning_value)
y$binned_median <- rep(rollapply(y$median, width = binning_value, by = binning_value,
FUN = mean, align = "left"), each=binning_value)
# Uses a plotting function for a single plotted actogram if Double_Single <- 'SP' is set by user
if (Double_Single == 'SP') {
ggplot(y, aes(Dec_time, y=get(mean_or_median_column), ymax=get(mean_or_median_column),
ymin=min(get(mean_or_median_column)))) +
geom_ribbon() +
facet_grid(date ~ .)+
geom_ribbon(fill="#0D226E") +
labs(title= x, x= "", y = "Counts/recording frequency")+
theme_bw()+
scale_x_continuous(breaks = c(0, 360,720, 1080, (1440-data_recording_frequency)),
labels=c("0 h", "6 h", "12 h", "18 h", "0 h")) +
theme(axis.text=element_text(size=14))+
theme(text = element_text(size=16))+
theme(plot.title = element_text(size = rel(2), hjust=0.5))+
coord_cartesian(ylim=c(0,as.numeric(ac_max_counts)))
} else {
# Uses a plotting function for a double plotted actogram if Double_Single <- 'DP' is set by the user
# Well, actually any value other than 'SP' would do it
# Thsese 3 lines double the actogram data
b <- arrange(y, date, Dec_time)
qqq <- lapply(unique(b$date), function(w) filter(b, date==w))
zr1 <- do.call("rbind", replicate(2, qqq, simplify = T))
# Generates a doubled dec time X scale
zr1$Dec_time_double <- c(rep(1:data_freq, length(unique(zr1$date))),
rep((data_freq+1):(data_freq*2), length(unique(zr1$date))))*data_recording_frequency
zr2 <- arrange(zr1, date, Dec_time_double)
zr3 <- zr2[(data_freq+1):nrow(zr2),] #drops the first repeated day
zr3 <- arrange(zr3, date, Dec_time_double)

#Generates another doubled dec time X scale after dropping the 1st day
zr3$Dec_time_double2 <- ((c(rep(1:(data_freq*2), length(unique(zr3$date))))[1:nrow(zr3)])* data_recording_frequency)

ggplot(zr3, aes(Dec_time_double2, y=get(mean_or_median_column), ymax=get(mean_or_median_column),
ymin=min(get(mean_or_median_column)))) +
theme_bw()+
facet_grid(date2 ~ .)+
geom_ribbon(fill="#0D226E") +
labs(title= x, x= "", y = "Number of counts/recording frequency")+
scale_x_continuous(breaks = c(1, 360,720, 1080, 1440, 1800, 2160, 2520, 2880),
labels=c("0h", "6 h", "12 h", "18 h", "0 h", "6 h", "12 h", "18 h", "0 h"))+
coord_cartesian(xlim=c(1,2880))+
theme(plot.title = element_text(size = rel(2), hjust=0.5))+
theme(axis.text=element_text(size=14))+
theme(text = element_text(size=16))+
coord_cartesian(ylim=c(0,as.numeric(ac_max_counts)))
}
}
})
# Assembles mean actograms using marrangeGrob
plot<-marrangeGrob(lapply(unique(mmf$Condition), function(x) FUN=mean_actogram(x)),
ncol=1, nrow = length(unique(mmf$Condition)), top ="")

ggsave("multipage.pdf", plot, width = 10, height = 80, units = "in",limitsize = FALSE)


```

## Mean periodograms
#Supplemental Figure15C
```{r, warning=FALSE, message=FALSE}

Mean_period_by_condition_rhythmic <- read.csv("/Volumes/NO NAME/Clk856DDN/Mean_periodograms_data.csv") #Reads a csv file

knitr::kable(head(Mean_period_by_condition_rhythmic)) #Displays the first 6 rows of a data frame
```
Order list 
```{r}

Mean_period_by_condition_rhythmic$Condition <-factor(Mean_period_by_condition_rhythmic$Condition,levels = c("CS_M","CS_F", "Trp_Fru_FLP_M","Trp_Fru_FLP_F","CLK_TRPA1_M","CLK_TRPA1_F"))
```



#### Column description:

* X - row index
* Condition - a user-defined name of an experimental condition
* Period - period tested by the Chi-Square algorithm [hours]
* Mean_Qp.act - Mean Chi-Square Qp.act value across all rhythmic individuals in a condition
* SEM_Qp.act - SEM of the Mean_Qp.act
* N_of_rhythmic_flies - number of rhythmic flies passing the Qp.act/Qp.sig threshold
* Mean_Qp.sig - Chi-Square Qp.sig period significance threshold
* Mean_Qp.act_Qp.sig_ratio - Ratio of Mean_Qp.act / Mean_Qp.sig. Values used for calling the strongest peak.
* SEM_Qp.act_Qp.sig_ratio - SEM of the Mean_Qp.act_Qp.sig_ratio
\newline

```{r, fig.height = 5, fig.width = 9, fig.align = 'center', warning=FALSE, message=FALSE}
#Overlapping mean periodograms

 Mean_periodograms_plot_overlapping <- ggplot(Mean_period_by_condition_rhythmic, 
                                              aes(x=Period, y=Mean_Qp.act, colour = Condition)) + 
      geom_point()+
      geom_line()+
      geom_line(data=Mean_period_by_condition_rhythmic, aes(x=Period, y = Mean_Qp.sig), size=0.5, colour="black")+
      theme_bw()+
      labs(x="Circadian Period [h]", y="Mean Qp.act")+
      theme(plot.title = element_text(size = rel(2), hjust=0.5))+
      scale_x_continuous(breaks = seq(4, 56, 2)) +
      geom_vline(xintercept = 24, linetype = 2 ) +
      theme(legend.text=element_text(size=18))+
      theme(legend.title = element_text(size=18))+
      theme(axis.text.x=element_text(hjust=0.5, size=15, face="bold"))+
      theme(axis.text.y=element_text(hjust=0.5, size=15, face="bold"))+
      theme(axis.title=element_text(size=18))+
      guides(colour = guide_legend(override.aes = list(size=1))) + 
      geom_ribbon(aes(ymin=Mean_Qp.act-SEM_Qp.act, ymax=Mean_Qp.act+SEM_Qp.act), linetype=1, alpha=0.01)
    
      #scale_colour_manual(values=c("blue", "red")) # Add this line to customize colors
 
 
Mean_periodograms_plot_overlapping

```
## Period peaks
#Supplemental Figure15D
```{r, warning=FALSE, message=FALSE}

Indivdual_fly_period_peaks_data <- read.csv("/Volumes/NO NAME/Clk856DDN/Individual_fly_period_peaks.csv")

knitr::kable(head(Indivdual_fly_period_peaks_data))  
  
```



Order list 
```{r}

Indivdual_fly_period_peaks_data$Condition <-factor(Indivdual_fly_period_peaks_data$Condition,levels = c("CS_M","CS_F", "Trp_Fru_FLP_M","Trp_Fru_FLP_F","CLK_TRPA1_M","CLK_TRPA1_F"))
```


#### Column description:
* X - row index
* Condition - a user-defined name of an experimental condition
* channel - DAM system channel
* Period - Chi-Square Period peak [hours] of an individual
* Qp.act - Chi-Square Qp.act period peak value of an individual
* Qp.sig - Chi-Square Qp.sig period peak significance threshold
* Act_Sig_ratio - Qp.act/Qp.sig ratio. Provides information about the period strength 
\newline

```{r, fig.height = 5, fig.width = 5, fig.align = 'center', warning=FALSE, message=FALSE}

# Period peaks boxplot

 ind_periods<- ggplot(na.omit(Indivdual_fly_period_peaks_data), aes(x=Condition, y=Period, colour = Condition)) +
      geom_boxplot(alpha=0.7)+
      geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.5, aes(fill=Condition))+
      labs(y="Circadian Period [h]", x="") +  #adds/removes axis lables
      theme(legend.title=element_blank())+ #removes legend title
      theme_bw()+
      theme(axis.text.x=element_text(angle=50, hjust=1, size=18, face="bold"))+
      theme(axis.text.y=element_text(hjust=1, size=16, face="bold"))+
      theme(axis.title.y = element_text(color="black", size=18))+         #axis title
      theme(legend.title = element_blank())+
      theme(legend.text = element_text(size=18))+
      labs(title= "Circadian period peaks")+
      theme(plot.title = element_text(size = rel(2), hjust=0.5))+
      theme(legend.position="none")

ind_periods
```


#Supplemental Figure15E
```{r, fig.height = 5, fig.width = 5, fig.align = 'center', warning=FALSE, message=FALSE}

# Box plot of individual Period strengths  

    ind_strength<- ggplot(na.omit(Indivdual_fly_period_peaks_data), aes(x=Condition, y=Act_Sig_ratio, colour = Condition)) +
      geom_boxplot(alpha=0.7)+
      geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.5, aes(fill=Condition))+
      labs(y="Qp.act/Qp.sig", x="") +  #adds/removes axis lables
      theme(legend.title=element_blank())+ #removes legend title
      theme_bw()+
      theme(axis.text.x=element_text(angle=50, hjust=1, size=18, face="bold"))+
      theme(axis.text.y=element_text(hjust=1, size=16, face="bold"))+
      theme(axis.title.y = element_text(color="black", size=18))+         #axis title
      theme(legend.title = element_blank())+
      theme(legend.text = element_text(size=18))+
      labs(title= "Circadian period strength")+
      theme(plot.title = element_text(size = rel(2), hjust=0.5))+
      theme(legend.position="none")
    
    ind_strength

```

